---
title: "Group_11_Analysis"
author: "Group11"
date: "2023-03-09"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Aims of the Analysis

# Exploratory Data Analysis

```{r Loading, message=FALSE, warning=FALSE}
# load library
library(tidyr)
library(ggplot2)
library(skimr)
library(dplyr)
library(gridExtra)
library(GGally)
library(gmodels)
library(stats)
library(sjPlot)
library(jtools)
library(MASS)

# load the dataset for group 11
coffee <- read.csv('https://raw.githubusercontent.com/rrachelxi/DAS-Group-11/main/dataset11.csv')
```


```{r skim data}
# skim the dataset
skim_without_charts(coffee)

# There are missing values in altitude_mean_meters, country_of_origin and harvested
# For altitude_mean_meters:
# missing values are replaced with altitude_mean_meters mean values
# For country_of_origin and harvested:
# since observations with NA are in relatively small amount, they are dropped
```

```{r deal with missing values}
# Calculate the mean value for altitude_mean_meters
mean_altitude <- mean(coffee$altitude_mean_meters,na.rm = TRUE)

# Perform data wrangling to the dataset
coffee_m <- coffee %>%
  mutate(
    # turn harvested into the factor type  for analysis
    harvested = as.factor(harvested), 
    # turn country_of_origin into the factor type for analysis
    country_of_origin = as.factor(country_of_origin), 
    # turn Qualityclass into the factor type for analysis
    Qualityclass = as.factor(Qualityclass), 
    # Replace NA in altitude_mean_meters by the mean value)
    altitude_mean_meters = replace_na(altitude_mean_meters, mean_altitude)) %>% 
  filter(
    !is.na(country_of_origin) & # drop observations with NA in country_of_origin
    !is.na(harvested)) # drop observations with NA in harvested

# Skim the dataset the second time to check if NA values still present
skim_without_charts(coffee_m)

# The dataset is now with no NA values
```


```{r pair plot of coffee_m, message=FALSE, warning=FALSE, fig.height = 10 , fig.width = 12, fit.align = "center", fig.cap = "\\label{fig:pp1} Pair plot of numeric variables and Qualityclass."}
# Draw pair plot to visualize data
ggpairs(coffee_m[,c(-1,-7)], 
        aes(color=Qualityclass,alpha=0.2), 
        title = "Distribution between variables")

# Outliers are spotted for aroma, acidity and altitude_mean_meters
```

```{r scatter plot aroma acidity, fig.cap = "\\label{fig:sp1} Dsitribution between aroma and acidity."}
# Draw scatter plots to show outliers more clearly
# Draw scatter plots for aroma and acidity
ggplot(coffee_m,aes(x=aroma,y=acidity,color=Qualityclass)) +
  geom_point() +
  geom_jitter(width = 0.2, height = 0.2) +
  labs(x = "Aroma grade (ranging from 0-10)", 
       y = "Acidity grade (ranging from 0-10)",
       title = "Aroma grade and acidity grade for coffee quality")
```
#It can be noticed good quality coffee bean tend to have a higher level of acidity and aroma,comparing with the poor quality coffee bean.

```{r scatter plot altitude_mean_meters, fig.cap = "\\label{fig:sp2} Scatter plot of altitude_mean_meters."}
# Draw scatter plots for altitude_mean_meters
ggplot(coffee_m,aes(x=Qualityclass,y=altitude_mean_meters,color=Qualityclass)) +
  geom_point() +
  geom_jitter(width = 0.2, height = 0.2) +
  labs(x = "Quality Class", 
       y = "Mean altitude of the growers farm (in metres)",
       title = "Mean altitude of the growers group by coffee quality") +
  theme(legend.position = 'none')
  
```

```{r drop outliers}
# Drop outlier observations from the dataset
coffee_w <- coffee_m %>%
  filter(
    aroma > 6 & 
    acidity > 6 & 
    altitude_mean_meters < 100000)

# Skim the dataset the third time to give insights on variables
skim_without_charts(coffee_w)
```

```{r pair plot of coffee_w, fig.height = 10 , fig.width = 12, fit.align = "center", fig.cap = "\\label{fig:pp2} Pair plot of numeric variables and Qualityclass (without outliers).", message=FALSE, warning=FALSE}
# For numeric variables, pair plot, inclduing boxplots, is drawn to show distribution and correlation
ggpairs(coffee_w[c(-1,-7)], 
        aes(color=Qualityclass,alpha=0.2), 
        title = "Distribution between variables(without outliers)")

# Collinearity is shown in pair plot, could be discussed in extension/further work
```
#There is a relatively strong correlation between aroma and flavor (correlation:0.770),similar as acidity and aroma(correlation:0.643).

```{r barplot of harvested, fig.cap = "\\label{fig:bp1} Bar plot of quality class by harvested year."}
# For categorical variables
# Draw a barplot to visualize the distribution of harvested and Qualityclass
ggplot(coffee_w, aes(x=harvested, y= after_stat(prop), group=Qualityclass, fill=Qualityclass )) +
  geom_bar(position = "dodge") +
  labs(x = "Year the batch was harvested",
       y = "Proportion",
       title = "Porportions of quality class by year the batch was harvested")
```


```{r patble harvested} 
# Build a table of proportions to support the barplot of harvested and Qualityclass
prop.table(table(coffee_w$Qualityclass,coffee_w$harvested),2) %>%
  round(digits=2)
```
#The batch of year 2012 has the highest proportion of good and poor quality coffee, with the batch of year 2014 following behind.

```{r barplot of country,fig.cap = "\\label{fig:bp2} Barplot of quality class by country of origin."}
# Draw a barplot to visualize the distribution of country_of_origin and Qualityclass
ggplot(coffee_w, aes(x=country_of_origin, group=Qualityclass )) +
  geom_bar(aes(y=after_stat(prop), fill = Qualityclass),  position = "dodge") +
  labs(x = "Country of origin",
       y = "Proportion",
       title = "Porportions of quality class by country of origin") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

```
#Mexico has the highest proportion of poor quality coffee bean(0.74) while Colombia has the highest proportion of good quality coffee bean(0.82).

```{r patble country_of_origin} 
# Build a table of proportions to support the barplot of country_of_origin and Qualityclass
prop.table(table(coffee_w$Qualityclass,coffee_w$country_of_origin),2) %>%
  round(digits=2)


```

# Statistical Modelling

```{r feature selection}
# Build the full glm model
model_full <- glm(
  Qualityclass ~ 
               aroma + 
               flavor +
               acidity +
               category_two_defects +
               altitude_mean_meters +
               harvested +
               country_of_origin,
  data = coffee_w,
  family = binomial(link="logit"))

# Conduct stepwise feature selection by AIC
model_step<-model_full %>%
  stepAIC(trace=TRUE)
```
#Based on AIC, the model with variable of aroma, flavor, acidity, category_two_defects, altitude_mean_meters and country_of_region would be chosen.  

```{r summary model_step}
# Pull summary of model chosen by stepwise selection
summary(model_step)
```

```{r model_step accuracy}
# Adding predict results from model_step to dataset
coffee_step <- coffee_w %>%
  mutate(logodds_poor = predict(model_step),
         probs_poor = fitted(model_step)) %>%
  mutate(odd_poor = exp(logodds_poor),
         class_pred = ifelse(probs_poor>0.5,"Poor","Good"))

# Check accuracy for model_step
sum(coffee_step$class_pred == coffee_step$Qualityclass)/nrow(coffee_step)
```
#The p value of aroma, flavor and acidity is  smaller than 0.05, which indicate these variables have a significant relationship with quality class.

```{r model with significant coefficients}
# Build a model with major terms with significant coefficients from model_step
model_1 <- glm(
  Qualityclass ~ 
               aroma + 
               flavor +
               acidity,
  data = coffee_w,
  family = binomial(link="logit"))

# Pull summary of model_1
summ(model_1)
```

```{r model_1 accuracy}
# Adding predict results from model_1 to dataset
coffee_1 <- coffee_w %>%
  mutate(logodds_poor = predict(model_1),
         probs_poor = fitted(model_1)) %>%
  mutate(odd_poor = exp(logodds_poor),
         class_pred = ifelse(probs_poor>0.5,"Poor","Good"))

# Check accuracy for model_1
sum(coffee_1$class_pred == coffee_1$Qualityclass)/nrow(coffee_1)
```
#model_1 has slightly higher AIC and less accuracy than model_step, but all coefficients are significant, thus model_1 is chosen as the final model.

```{r Odds (Poor quality), fig.cap = "\\label{fig:lo} Point estimates and confidence intervals for odds."}
# Plot point estimate and confidence intervals of model_1
plot_model(model_1, show.values = TRUE, 
           title = "Odds (Poor quality)",
           show.p = FALSE, digits=3)
```

```{r aroma, fig.cap = "\\label{fig:aroma} Probability of being poor quality by aroma."}
# Plot estimate possibilities by each explanatory variables in model_1
plot_model(model_1, type="pred", terms="aroma [all]", 
           axis.title = "Probability",
           title = "Probability of being poor quality by aroma")
```
#Based on model_1, probability of being poor quality would decrease with the higher level of aroma.

```{r flavor, fig.cap = "\\label{fig:flavor} Probability of being poor quality by flavor."}
# Plot estimate possibilities by each explanatory variables in model_1
plot_model(model_1, type="pred", terms="flavor [all]", 
           axis.title = "Probability",
           title = "Probability of being poor quality by flavor")
```
#Based on model_1, probability of being poor quality would decrease with the higher level of flavor.

```{r acidity, fig.cap = "\\label{fig:acidity} Probability of being poor quality by acidity"}
# Plot estimate possibilities by each explanatory variables in model_1
plot_model(model_1, type="pred", terms="acidity [all]", 
           axis.title = "Probability",
           title = "Probability of being poor quality by acidity")
```
#Based on model_1, probability of being poor quality would decrease with the higher level of acidity.

Summary: Based on model_1,since the p value of three variables are all smaller than 0.05, they are significantly related to the qualityclass of coffee bean.The coffee bean with higher level of aroma, flavor and acidity tends to have a higher probability of being good quality. In addition to this, other variable outside this model such as origin and harvested batch will also influence the probability of being good quaility, but they are not significant as these three variables above.
